---
title: "2019HW_week8_9_bangladesh_brms"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggthemes)
library(brms)
library(brmstools)
theme_set(theme_tufte(base_family = 'sans'))
options(mc.cores = parallel::detectCores())

data('bangladesh', package = 'rethinking')
d <- bangladesh; rm(bangladesh)
d <- d %>% 
  mutate(did = district %>% as_factor())
```
```{r}
get_prior(data=d, family = binomial,
            use.contraception ~ 0 + factor(did))
```
### Model 1: traditional fixed effects model
```{r, message=FALSE, cache=TRUE, results="hide"}
b_1 <- brm(data=d, family = binomial,
            use.contraception ~ 0 + factor(did),
            prior = prior(normal(0,1.5), class = b)
            )
```
```{r}
fixef(b_1) %>% 
  as.tibble() %>% 
  inv_logit_scaled() %>% 
  bind_cols(d %>% 
    group_by(did) %>% 
    summarise(sample_mean = mean(use.contraception)) 
    ) %>% 
  ggplot(aes(y = Estimate, ymax = `Q97.5`, ymin = `Q2.5`, x = did))+
  geom_errorbar()+
  geom_point(aes(y=sample_mean), shape = 1)
```

### Model 2: centered partial pooling model
```{r, cache=T, message=F, results='hide'}
b_2 <- brm(data = d, family = bernoulli,
           use.contraception ~ 1 + (1 | did),
           prior = prior(normal(0,1.5), class  = Intercept) + 
             prior(exponential(1), class = sd)
           )
```
```{r}
summary(b_2)
```

```{r}
b_1 <- add_criterion(b_1, "waic")
b_2 <- add_criterion(b_2, "waic")

w <- loo_compare(b_1, b_2, criterion = "waic")

print(w, simplify = F)
```
```{r}
coef(b_2, robust = T)$did %>% 
  inv_logit_scaled() %>% 
  as_tibble() %>% 
  bind_cols(d %>% 
    group_by(did) %>% 
    summarise(sample_mean = mean(use.contraception)) 
    ) %>%
  ggplot(aes(y = Estimate.Intercept, ymax = `Q97.5.Intercept`, ymin = `Q2.5.Intercept`, x = did))+
  geom_errorbar()+
  geom_point(aes(y=sample_mean), shape = 1)
```

### Model 5 varying slope with covariance

```{r}
get_prior(data = d, family = bernoulli,
          use.contraception ~ 1 + urban + (1 + urban | did)
          )
```


```{r, cache=T, message=F, results='hide'}
b_5 <- brm(data = d, family = bernoulli,
          use.contraception ~ 1 + urban + (1 + urban | did),
          prior = prior(normal(0,1.5), class  = Intercept) + 
                  prior(normal(0, 0.5), class = b) +
                  prior(exponential(1), class = sd) + 
                  prior(lkj(2), class = cor)
          )
```
```{r}
summary(b_5)
```

```{r}
coef(b_5, robust = T)$did[,,2]
```
```{r}
d %>% 
  ggplot(aes(x = as_factor(urban), y = use.contraception))+
  # geom_line()+
  geom_jitter(shape=1, height = 0)+
  facet_wrap(~did, ncol = 10)+
  # theme_fivethirtyeight()+
  theme(panel.grid.major.x = element_blank())
```
```{r}
predictions <- d %>% 
  modelr::data_grid(did, urban) %>% 
  tidybayes::add_predicted_draws(b_5) %>%
  # mutate(.prediction = inv_logit_scaled(.prediction)) %>% # do we need this or not?
  group_by(did, urban) %>% 
  summarize(estimate = mean(.prediction))
```
```{r}
d %>% 
  ggplot(aes(x = as_factor(urban), y = use.contraception))+
  geom_jitter(shape=1, height = 0)+
  facet_wrap(~did, ncol = 10)+
  geom_line(data = predictions,
            aes(x = as_factor(urban), y = estimate, group = did)
            )+
  theme(panel.grid.major.x = element_blank())
```
This is the second time I made this chart, and I think the first time I did it wrong by using the inv_logit to transform what came out of the link function. I think the predicted_draws and fitted_draws of brms have already been transformed. same convention for rethinking package.

```{r}
pp_check(b_5)
```

```{r}
d %>% 
  modelr::data_grid(did, urban) %>% 
  tidybayes::add_predicted_draws(b_5)
```
```{r}
d %>% 
  modelr::data_grid(did, urban) %>% 
  posterior_linpred(b_5, newdata = ., summary = F) %>% 
  range()
```

  fitted(b_5, transform = logit_scaled) %>% head
```{r}
```

```{r}
coefplot(b_5)
```
```{r}
tidyfitted(b_5)
```
```{r}
panels(b_5, xvar = "urban", data = T)
```

```{r}
spaghetti(b_5, xvar = "urban")
```

### Model 6 using matrix instead of slopes
```{r}
get_prior(data = d, family = bernoulli,
          use.contraception ~ 0 + factor(urban) + (0 + factor(urban) | did)
          )
```


```{r, cache=T, message=F, results='hide'}
b_6 <- brm(data = d, family = bernoulli,
          use.contraception ~ 0 + factor(urban) + (0 + factor(urban) | did),
          prior = prior(normal(0, 1.5), class = b) +
                  prior(exponential(1), class = sd) + 
                  prior(lkj(2), class = cor)
          )
```
```{r}
summary(b_6)
```

### Model 7 causal inference of age and living children

```{r, cache=T, message=F, results='hide'}
d <- d %>% 
  mutate(A = age.centered %>% scale,
         C = living.children %>% scale
         )
b_7 <- brm(data = d, family = bernoulli,
          use.contraception ~ 1 + urban + (1 + urban | did) + A + C,
          prior = prior(normal(0,1.5), class  = Intercept) + 
                  prior(normal(0, 0.5), class = b) +
                  prior(exponential(1), class = sd) + 
                  prior(lkj(2), class = cor)
          )
```

```{r}
summary(b_7)
```
### Model 9 total influence of age

```{r, cache=T, message=F, results='hide'}
b_9 <- brm(data = d, family = bernoulli,
          use.contraception ~ 1 + urban + (1 + urban | did) + A,
          prior = prior(normal(0,1.5), class  = Intercept) + 
                  prior(normal(0, 0.5), class = b) +
                  prior(exponential(1), class = sd) + 
                  prior(lkj(2), class = cor)
          )
summary(b_9)
```

### Model 8: influence of Age with interaction from number of children

```{r, cache=T, message=F, results='hide'}
b_8 <- brm(data = d, family = bernoulli,
          use.contraception ~ 1 + urban + (1 + urban | did) + A + A:C,
          prior = prior(normal(0,1.5), class  = Intercept) + 
                  prior(normal(0, 0.5), class = b) +
                  prior(exponential(1), class = sd) + 
                  prior(lkj(2), class = cor)
          )
summary(b_8)
```
 A interaction model cannot block the backdoor!

### Model 10: influence of Age when the number of living children is treated as montonic

```{r, cache=T, message=F, results='hide'}
b_10 <- brm(data = d, family = bernoulli,
          use.contraception ~ 1 + urban + (1 + urban | did) + A + mo(living.children),
          prior = prior(normal(0,1.5), class  = Intercept) + 
                  prior(normal(0, 0.5), class = b) +
                  prior(exponential(1), class = sd) + 
                  prior(lkj(2), class = cor)
          )
summary(b_10)
```
```{r}
marginal_effects(b_10)
```

