---
title: "2019HW_week8&9_bangladesh"
output: 
  github_document:
    toc: true
  always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggthemes)
library(rethinking)

data("bangladesh")
d <- bangladesh
d %>% glimpse

d <- d %>% 
  mutate(district_id = district %>% as.factor %>% as.integer)

dat <- with(d,
            list(dist_id = district_id,
                 contraception = use.contraception %>% as.integer()
                   )
            )

summarise_samples <- function(data, point_func = mean, range_func = HPDI, ...){
  point <- data.frame(
    estimate = apply(data, 2, point_func)
  )
  range_result <- apply(data, 2, range_func, ...) %>% t()
  colnames(range_result)<-c("lower", "upper")
  cbind(point, range_result)
}
```

## Bangladeshy birth rates

Would prefer to tackle the series of questions on bangladeshy birth rates questions first

### Load data

The professor is super kind to alert us the district is not a factor and actually contains one missing district: 54

I shall try to find it out

```{r missing}
rng <- range(d$district)
which(!(rng[1]:rng[2] %in% d$district))

```
and it worked!

### Model 1: traditional fixed effects model

should use better priors than norm(0,1)

```{r}
a <- rnorm(1e4,0,1.5)
p <- inv_logit(a)
dens(p)
```


```{r, cache=TRUE, message=F}
m_contra_1 <- ulam(
  alist(
    contraception ~ dbinom(1, p),
    logit(p) <- a[dist_id],
    a[dist_id] ~ dnorm(0,1.5)
  ), data = dat, cores = 4, chains = 4, log_lik = 4 
)
```
just a quick look at the posterier samples
```{r}
precis(m_contra_1, depth = 2) %>% data.frame -> post_summary
post_summary <- post_summary %>% 
  mutate(dist_id = 1:n(),
         ymin = X5.5. %>% inv_logit(),
         ymax = X94.5.%>% inv_logit(),
         y = mean %>% inv_logit(),
         pooling = "no"
         )
data_proportion <- d %>% group_by(district_id) %>% summarise(proportion = mean(use.contraception))
```

```{r}
post_summary %>%
  ggplot(aes(x = dist_id, y = y, ymax = ymax, ymin = ymin))+
  geom_point(shape = 1)+
  geom_linerange()+
  geom_point(data = data_proportion, aes(x = district_id, y = proportion), color = "red", inherit.aes = F)+
  labs(x = "district ID", y = "probability of using contraception")+
  theme_fivethirtyeight()
  
```

### Model 2 Centered partial pooling version

Centered version first and see how it goes
```{r, cache=TRUE, message=F}
m_contra_2 <- ulam(
  alist(
    contraception ~ dbinom(1, p),
    logit(p) <- a[dist_id],
    a[dist_id] ~ dnorm(mu, sigma),
    mu ~ dnorm(0, 1.5),
    sigma ~ dexp(1)
  ), data = dat, cores = 4, chains = 4, log_lik = 4 
)
```


```{r}
precis(m_contra_2, depth = 2) %>% plot
```
```{r}
precis(m_contra_2, depth = 2) %>% data.frame -> post_summary2
post_summary2 <- post_summary2 %>% 
  mutate(dist_id = 1:n()+0.3,
         ymin = X5.5. %>% inv_logit(),
         ymax = X94.5.%>% inv_logit(),
         y = mean %>% inv_logit(),
         pooling = "partial"
         )
data_proportion <- d %>% group_by(district_id) %>% summarise(proportion = mean(use.contraception))
```

```{r}
post_summary2 %>%
  ggplot(aes(x = dist_id, y = y, ymax = ymax, ymin = ymin))+
  geom_point(shape = 1)+
  geom_linerange()+
  geom_point(data = data_proportion, aes(x = district_id, y = proportion), color = "red", inherit.aes = F)+
  labs(x = "district ID", y = "probability of using contraception")+
  theme_fivethirtyeight()
  
```
Time to contrast the models
```{r}
post_summary <- post_summary %>% 
  bind_rows(post_summary2)
```
```{r}
post_summary %>%
  ggplot(aes(x = dist_id, y = y, ymax = ymax, ymin = ymin, color = pooling))+
  geom_point(shape = 1, alpha=0.5)+
  geom_linerange(alpha=0.5)+
  geom_point(data = data_proportion, aes(x = district_id, y = proportion), color = "red", inherit.aes = F)+
  geom_hline(yintercept = mean(d$use.contraception), linetype = "dashed")+
  labs(x = "district ID", y = "probability of using contraception")+
  theme_fivethirtyeight()
```

This is problemmatic - if we look at district 3, the partial pooling results seems to gravitate towards below 0.5, where data is clearly 1. checked the solutions and yes indeed it should shrink to the global mean!

```{r}
d %>% 
  count(district) %>% 
  filter(n<10)

```


### Model 3 non-centered version of the hierachical model

```{r, cache=TRUE, message=F}
m_contra_3 <- ulam(
  alist(
    contraception ~ bernoulli(p),
    logit(p) <- a_bar + a[dist_id] * sigma_dist,
    a_bar ~ dnorm(0, 1.5),
    sigma_dist ~ dexp(1),
    a[dist_id]~dnorm(0,1)
  ),
  data = dat, cores = 4, chains = 4
)
```
```{r}
precis(m_contra_3, depth = 2) %>% plot

post_summary3 <- precis(m_contra_3, depth = 2)
```
### Model 4 brms version of the non-centered

instead of trying to re-create the parameter and compare with the centered version, decided to use brms to checking.

before loading the brms, maybe I need to unload the rethinking package first?

```{r}
library(brms)
```
```{r}
get_prior(data = dat, family = bernoulli,
                  contraception ~  (1 | dist_id))
```

Cannot test this as brms crashes rstudio if I try to fit model more than 1 layer

```{r, message=F, cache=TRUE}
b_contra_3 <- brm(data = dat, family = bernoulli,
                  contraception ~ 1 + (1 | dist_id),
                  prior = prior(normal(0,1.5), class = Intercept) +
                          prior(exponential(1), class = sd)
                  )
```
```{r}
b_contra_3 %>% summary
```

```{r}
precis(m_contra_3)
```
```{r}
coef(b_contra_3)$dist_id[1:5,,]
```
```{r}
precis(m_contra_2, depth = 2)
```
basically the brms coef function returns the coefficient of variables as if it is a centered model.
```{r}
fixef(b_contra_3)
```
```{r}
precis(m_contra_3, depth = 2)
```
```{r}
ranef(b_contra_3)$dist_id[1:3,,]
```


### Model 5 varying slope with covariance

```{r}
dat <- with(d,
            list(dist_id = district_id,
                 contraception = use.contraception %>% as.integer(),
                 urban = urban %>% as.integer()
                   )
            )
```

```{r, message=F, cache=T}
m_contra_5 <- ulam(
  alist(
    contraception ~ bernoulli(p),
    logit(p) <- a[dist_id] + b[dist_id] * urban,
    c(a, b)[dist_id] ~ multi_normal(c(mu_a, mu_b), Rho, Sigma),
    mu_a ~ dnorm(0, 1.5),
    mu_b ~ dnorm(0, 0.5),
    Sigma ~ dexp(1),
    Rho ~ lkj_corr(2)
  ),
  data = dat, core = 4, chains = 4
)
```
```{r}
precis(m_contra_5, depth = 3) 
```
The correlation between $a$ and $b$ is negative. This does not mean being in the urban area depresses the use of contraception, that would be counter intuitive. In fact, $mu_b$ is definitely positive, which showed that the effect of being in an urban area, everything else being equal, is more conducive to the adoption of contraceptives.

What the correlation between $a$ and $b$ showed is that for a new dsitrict, the higher the expected adoption rate, the lower the impact of that new district being urban. Put it in story form, if a district is prosperous, they it is more likely the women in the district would use contraception, and if the adoption rate of contraception is high already, the difference between urban and rural area would be smaller.

```{r, fig.width=15}
p <- d%>% 
  group_by(district, urban) %>% 
  summarise(proportion = mean(use.contraception)) %>% 
  ungroup() %>% 
  mutate(x = 1:n()) %>% 
  ggplot(aes(x = x, y = proportion, group = district, text = paste(district, "urban =", urban)))+
  geom_point(shape=1)+
  geom_line()+
  # facet_wrap(~district_id)+
  # labs(x = "urban = 1", y = "contraception = 1")+
  theme_fivethirtyeight()
p
```
```{r}
d_summary <- d%>% 
  group_by(district, urban) %>% 
  summarise(proportion = mean(use.contraception)) %>% 
  ungroup() %>% 
  mutate(x = 1:n())
```
```{r}
p <- d %>% 
  ggplot(aes(x = as_factor(urban), y = use.contraception))+
  # geom_line()+
  geom_jitter(shape=1, height = 0)+
  facet_wrap(~district_id, ncol = 10)+
  theme_fivethirtyeight()+
  theme(panel.grid.major.x = element_blank())

p
```

```{r}
data_grid <- d %>% 
  modelr::data_grid(district_id, urban) %>% 
  rename(dist_id = district_id)
result <- link(m_contra_5, data=data_grid) %>% 
  apply(2, inv_logit) %>% 
  summarise_samples()

simulated <- cbind(data_grid, result) 
```

```{r, warning=FALSE}
d_2 <- d %>% 
  rename(dist_id = district_id)
simulated %>% 
  ggplot(aes(x = as_factor(urban), y = estimate))+
  geom_jitter(data = d_2, aes(y = use.contraception), shape=1, height = 0)+
  geom_point()+
  geom_errorbar(aes(ymax = upper, ymin = lower))+
  geom_line(aes(group = dist_id))+
  facet_wrap(~dist_id, ncol = 10)+
  theme_fivethirtyeight()+
  theme(panel.grid.major.x = element_blank())

```

```{r}
data_grid <- d %>% 
  modelr::data_grid(district_id, urban) %>% 
  rename(dist_id = district_id)
result <- sim(m_contra_5, data=data_grid) %>% 
  apply(2, inv_logit) %>% 
  summarise_samples()

simulated <- cbind(data_grid, result) 

simulated %>% 
  ggplot(aes(x = as_factor(urban), y = estimate))+
  geom_jitter(data = d_2, aes(y = use.contraception), shape=1, height = 0)+
  geom_point()+
  geom_errorbar(aes(ymax = upper, ymin = lower))+
  geom_line(aes(group = dist_id))+
  facet_wrap(~dist_id, ncol = 10)+
  theme_fivethirtyeight()+
  theme(panel.grid.major.x = element_blank())

```

```{r}
sample_trans <- d_2 %>% 
  group_by(dist_id, urban) %>% 
  summarise(proportion = mean(use.contraception)) %>% 
  ungroup() %>% 
  mutate(area = ifelse(urban == 0, "rural", "urban")) %>%
  select(-urban) %>% 
  group_by(dist_id) %>% 
  spread(area, proportion) %>% 
  replace(is.na(.), 0) %>% 
  mutate(type = "sample")
```

```{r}
sample_trans %>% 
  ggplot(aes(urban, rural))+
  geom_point(shape = 1)+
  geom_abline(slope = 1, linetype = "dashed")+
  theme_fivethirtyeight()+
  labs(x = "urban", y = "rural")
  
```

```{r}
sim_trans <- simulated %>%
  select(-lower, -upper) %>%
  rename(proportion = estimate) %>%
  mutate(area = ifelse(urban == 0, "rural", "urban")) %>%
  select(-urban) %>% 
  group_by(dist_id) %>% 
  spread(area, proportion) %>% 
  mutate(type = "sim")
  
```

```{r}
bind_rows(sample_trans, sim_trans) %>% 
  ggplot(aes(urban, rural))+
  geom_point(aes(shape = type))+
  geom_line(aes(group = dist_id), alpha=1/5)+
  geom_abline(slope = 1, linetype = "dashed")+
  theme_fivethirtyeight()+
  labs(x = "urban", y = "rural")
```
### Model 6 using matrix instead of slopes

Prof proposed this model after previous model shown that the slope is only neccessary to encode the negative correlation pattern.

```{r, cache=T, message=F}
dat$uid = dat$urban + 1L
m_contra_6 <- ulam(
  alist(
    contraception ~ bernoulli(p),
    logit(p) <- a[dist_id,uid],
    vector[2]:a[dist_id] ~ multi_normal(c(r_avg, u_avg), Rho, Sigma),
    c(r_avg, u_avg) ~ dnorm(0, 1.5),
    Rho ~ lkj_corr(2),
    Sigma ~ dexp(1)
  ), data = dat, cores = 4, chains = 4
)
```
```{r}
precis(m_contra_6, depth = 3)
```
### Model 7 causal inference of age and living children
```{r}
dat <- with(d,
            list(dist_id = district_id,
                 contraception = use.contraception %>% as.integer(),
                 urban = urban %>% as.integer(),
                 age = age.centered,
                 children = living.children
                )
            )
```

```{r, message=F, cache=T}
m_contra_7 <- ulam(
  alist(
    contraception ~ bernoulli(p),
    logit(p) <- a[dist_id] + b[dist_id] * urban + bAge * age + bC * children + bAC * age * children,
    c(a, b)[dist_id] ~ multi_normal(c(mu_a, mu_b), Rho, Sigma),
    mu_a ~ dnorm(0, 1.5),
    mu_b ~ dnorm(0, 0.5),
    c(bAge, bC, bAC) ~ dnorm(0, 0.5),
    Sigma ~ dexp(1),
    Rho ~ lkj_corr(2)
  ),
  data = dat, core = 4, chains = 4
)
```


```{r}
precis(m_contra_7, depth = 3)
```
the interaction model showed older women actually use contraception more, after the impact of the interaction with # of living children had been isolated. this makes more sense as it is more dangerous for older women to have children. 

### Model 8 causal inference of age and living children without interaction
```{r, message=F, cache=T}
dat_list <- with(dat,
                 list(
                   C = contraception,
                   did = dist_id,
                   urban = urban,
                   age = standardize(age),
                   children = standardize(children)
                 )
                 )
m_contra_8 <- ulam(
  alist(
    C ~ bernoulli(p),
    logit(p) <- a[did] + b[did] * urban + bAge * age + bK * children,
    c(a, b)[did] ~ multi_normal(c(mu_a, mu_b), Rho, Sigma),
    mu_a ~ dnorm(0, 1),
    mu_b ~ dnorm(0, 0.5),
    c(bAge, bK) ~ dnorm(0, 0.5),
    Sigma ~ dexp(1),
    Rho ~ lkj_corr(2)
  ),
  data = dat_list, core = 4, chains = 4
)
```
```{r}
precis(m_contra_8)
```
the model controlling for the number of children, showing older women tends not to use contraception

### Model 9 total influence of age
```{r, cache=T, message=F}
m_contra_9 <- ulam(
  alist(
    C ~ bernoulli(p),
    logit(p) <- a[did] + b[did] * urban + bAge * age,
    c(a, b)[did] ~ multi_normal(c(mu_a, mu_b), Rho, Sigma),
    mu_a ~ dnorm(0, 1),
    mu_b ~ dnorm(0, 0.5),
    bAge ~ dnorm(0, 0.5),
    Sigma ~ dexp(1),
    Rho ~ lkj_corr(2)
  ),
  data = dat_list, core = 4, chains = 4
)
```
```{r}
precis(m_contra_9)
```
the model without controlling for the number of children, showing older women tends to use contraception

### Model 10 influence of age with interaction
```{r, cache=T, message=F}
m_contra_10 <- ulam(
  alist(
    C ~ bernoulli(p),
    logit(p) <- a[did] + b[did] * urban + bAge * age + bK * children + bAK * age * children,
    c(a, b)[did] ~ multi_normal(c(mu_a, mu_b), Rho, Sigma),
    mu_a ~ dnorm(0, 1),
    mu_b ~ dnorm(0, 0.5),
    c(bAge, bK, bAK) ~ dnorm(0, 0.5),
    Sigma ~ dexp(1),
    Rho ~ lkj_corr(2)
  ),
  data = dat_list, core = 4, chains = 4
)
```
```{r}
precis(m_contra_10)
```
### Model 11 children as categorical predictor
```{r, message=F, cache=T}
dat_list <- with(dat,
                 list(
                   C = contraception,
                   did = dist_id %>% as.integer(),
                   urban = urban,
                   age = standardize(age),
                   children = children %>% as.integer(),
                   alpha = rep(2, 3)
                 )
                 )
m_contra_11 <- ulam(
  alist(
    C ~ bernoulli(p),
    logit(p) <- a[did] + b[did] * urban + bAge * age + bK * sum(delta_j[1:children]),
    c(a, b)[did] ~ multi_normal(c(mu_a, mu_b), Rho, Sigma),
    mu_a ~ dnorm(0, 1),
    mu_b ~ dnorm(0, 0.5),
    c(bAge, bK) ~ dnorm(0, 0.5),
    Sigma ~ dexp(1),
    Rho ~ lkj_corr(2),
    vector[4]: delta_j <<- append_row(0, delta),
    simplex[3]: delta ~ dirichlet(alpha)
  ),
  data = dat_list, core = 4, chains = 4
)
```
```{r}
precis(m_contra_11, depth = 2)
```
delta[1] is 0.74, hogging over 70% of the impact from having children, that means the women adopt contraception unevenly, few women take contraception after having 1 child, much more after 2 children, and such effects taper off after 3 or more children.
```{r}

```

