---
title: "Chapter 3 Practice"
output: html_notebook
---

```{r setup}
library(data.table)
library(rethinking)

```
This code will give you a specific set of samples, so that you can check your answers exactly.
```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
```

Use the values in samples to answer the questions that follow.
3E1. How much posterior probability lies below p = 0:2?

```{r}
mean(samples<0.2)
```
3E2. How much posterior probability lies above p = 0:8?
```{r}
mean(samples>0.8)
```
3E3. How much posterior probability lies between p = 0:2 and p = 0:8?
```{r}
mean(samples>=0.2 & samples<=0.8)
```
3E4. 20% of the posterior probability lies below which value of p?
```{r}
quantile(samples, 0.2)
```
3E5. 20% of the posterior probability lies above which value of p?
```{r}
quantile(samples, 0.8)
```
3E6. Which values of p contain the narrowest interval equal to 66% of the posterior probability?
```{r}
HPDI(samples, .66)
```
```{r}
HPDinterval <- function(obj, prob = 0.95, ...)
{
    obj <- as.matrix(obj)
    vals <- apply(obj, 2, sort)
    if (!is.matrix(vals)) stop("obj must have nsamp > 1")
    nsamp <- nrow(vals)
    npar <- ncol(vals)
    gap <- max(1, min(nsamp - 1, round(nsamp * prob)))
    init <- 1:(nsamp - gap)
    inds <- apply(vals[init + gap, ,drop=FALSE] - vals[init, ,drop=FALSE],
                  2, which.min)
    ans <- cbind(vals[cbind(inds, 1:npar)],
                 vals[cbind(inds + gap, 1:npar)])
    dimnames(ans) <- list(colnames(obj), c("lower", "upper"))
    attr(ans, "Probability") <- gap/nsamp
    ans
}
```
```{r}
HPDinterval(samples, 0.66)
```
3E7. Which values of p contain 66% of the posterior probability, assuming equal posterior probability
both below and above the interval?
```{r}
quantile(samples, c(.17, .83))
```
Medium.
3M1. Suppose the globe tossing data had turned out to be 8 water in 15 tosses. Construct the posterior
distribution, using grid approximation. Use the same flat prior as before.
```{r}
normalize <- function(variable){
  variable / sum(variable)
}
```

```{r}
p_grid <- seq(0,1, length.out = 1000)
prior <- rep(1, 1000)
likelihoods <- dbinom(8, 15, p_grid)

posterior <- normalize(prior * likelihoods)

```
3M2. Draw 10,000 samples from the grid approximation from above. Then use the samples to calculate
the 90% HPDI for p.
```{r}
samples <- sample(p_grid, 1e4, replace = T, prob = posterior)
HPDI(samples, 0.9)
```

3M3. Construct a posterior predictive check for this model and data. This means simulate the distribution
of samples, averaging over the posterior uncertainty in p. What is the probability of observing
8 water in 15 tosses?

Method 1 average over samples and calculate the pmf value
```{r}
dbinom(8, 15, prob = mean(samples))
```
Method 2, for each sample p, simulate N results of 15 tosses and calculate the percentage equal to 8
```{r}
sim <- sapply(samples, function(sample_i) rbinom(1e3, 15, prob = sample_i))
dim(sim)
mean(sim==8)
```
this is also method 2, as the samples in the below expression is recycled, the size of the rbinom is determined by n and n only
```{r}
sim <- rbinom(1e7, 15, prob = samples)
mean(sim==8)
```


3M4. Using the posterior distribution constructed from the new (8/15) data, now calculate the probability
of observing 6 water in 9 tosses.
```{r}
mean(sim==6)
```

3M5. Start over at 3M1, but now use a prior that is zero below p = 0:5 and a constant above p = 0:5.
This corresponds to prior information that a majority of the Earth’s surface is water. Repeat each
problem above and compare the inferences. What difference does the better prior make? If it helps,
compare inferences (using both priors) to the true value p = 0:7.
```{r}
p_grid <- seq(0,1, length.out = 1000)
prior <- ifelse(p_grid<0.5, 0, 1)
likelihoods <- dbinom(8, 15, prob = p_grid)

posterior <- normalize(prior * likelihoods)
```
```{r}
plot(p_grid, posterior)
```
```{r}
samples <- sample(p_grid, 1e4, prob = posterior, replace = T)
HPDI(samples, 0.9)
```
```{r}
sims <- rbinom(1e6, 15, prob = samples)
mean(sims==8)
mean(sims==6)

```
Hard.
Introduction. The practice problems here all use the data below. These data indicate the gender
(male=1, female=0) of officially reported first and second born children in 100 two-child families.
So for example, the first family in the data reported a boy (1) and then a girl (0). The second family
reported a girl (0) and then a boy (1). The third family reported two girls. You can load these two
vectors into R’s memory by typing:
```{r}
data(homeworkch3)
```
```{r}
sum(birth1 + birth2)
```

3H1. Using grid approximation, compute the posterior distribution for the probability of a birth
being a boy. Assume a uniform prior probability. Which parameter value maximizes the posterior
probability?

```{r}
#how many trials
trials <- length(birth1) + length(birth2)
# how many boys
boys <- sum(birth1 + birth2)
```

Assume independence between each child's gender
```{r}
p_grid <- seq(0, 1, length.out = 1000)

prior <- rep(1, 1000)

likelihoods <- dbinom(boys, trials, p_grid)

posterior <- normalize(prior * likelihoods)

plot(posterior ~ p_grid)
```
```{r}
p_grid[which.max(posterior)]
```


3H2. Using the sample function, draw 10,000 random parameter values from the posterior distribution
you calculated above. Use these samples to estimate the 50%, 89%, and 97% highest posterior
density intervals.
```{r}
p_samples <- sample(p_grid, 1e4, replace = T, prob = posterior)
HPDI(p_samples, c(.5, .89,.97))
```


3H3. Use rbinom to simulate 10,000 replicates of 200 births. You should end up with 10,000 numbers,
each one a count of boys out of 200 births. Compare the distribution of predicted numbers
of boys to the actual count in the data (111 boys out of 200 births). There are many good ways to
visualize the simulations, but the dens command (part of the rethinking package) is probably the
easiest way in this case. Does it look like the model fits the data well? That is, does the distribution
of predictions include the actual observation as a central, likely outcome?
```{r}
sims <- rbinom(1e4, trials, prob = p_samples)
dens(sims)
abline(v=111)
```


3H4. Now compare 10,000 counts of boys from 100 simulated first borns only to the number of boys
in the first births, birth1. How does the model look in this light?
```{r}
birth1_trials <- length(birth1)
sim_birth1 <- rbinom(1e4, birth1_trials, prob = p_samples)
dens(sim_birth1)
abline(v=sum(birth1))
```

3H5. The model assumes that sex of first and second births are independent. To check this assumption,
focus now on second births that followed female first borns. Compare 10,000 simulated counts
of boys to only those second births that followed girls. To do this correctly, you need to count the
number of first borns who were girls and simulate that many births, 10,000 times. Compare the
counts of boys in your simulations to the actual observed count of boys following girls. How does the
model look in this light? Any guesses what is going on in these data?
```{r}
births <- data.table(birth1=birth1, birth2=birth2)
```
```{r}
births[birth1 == 0, .(mean(birth2))]$V1
```

```{r}
trials <- births[birth1 == 0, .(len=length(birth2))]$len
boys <- births[birth1 == 0, .(boys=sum(birth2))]$boys
```
```{r}
sims <- rbinom(1e4, trials, prob = p_samples)
dens(sims, adj = .1)
abline(v = boys)
```


