---
title: "Chapter8"
output: 
  html_document: 
    fig_caption: yes
    fig_width: 10
    fig_height: 10
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# options(fig.width=10, fig.height=10)
options(device = Cairo::CairoWin)
# options(res = 720, dpi=700)
library(rethinking)
data(rugged)
d <- rugged

```
Make log version of response
```{r}
d$log_gdp = log(d$rgdppc_2000)
```
Filter out missing data
```{r}
dd <- d[complete.cases(d$rgdppc_2000),]
```
Rescale variables
```{r}
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)
```
Split countries between Africa and Non-Africa
```{r}
d.A0 <- dd[dd$cont_africa == 0 , ]
d.A1 <- dd[dd$cont_africa == 1 , ]
```
```{r}
m8.1 <- quap(
  alist(
    log_gdp_std ~ dnorm(mu, sigma),
    mu <- a + b * (rugged_std - 0.215),
    a ~ dnorm(1, 0.1),
    b ~ dnorm(0, 0.3),
    sigma ~ dexp(1)
  ), data = d.A1
)
```
Look at prior predictions of m8.1
```{r}
prior <- extract.prior(m8.1)

# set up the plot dimensions
plot( NULL , xlim=c(0,1) , ylim=c(0.5,1.5) ,
xlab="ruggedness" , ylab="log GDP" )

abline( h=min(dd$log_gdp_std) , lty=2 )
abline( h=max(dd$log_gdp_std) , lty=2 )

# Draw 50 lines from the prior
rugged_seq <- seq(-0.1, 1.1, length.out = 30)
mu <- link(m8.1, post = prior, data = list(rugged_std = rugged_seq))
for (i in 1:50) lines(rugged_seq, mu[i,], col = col.alpha("black", 0.3))
```
```{r}
m8.2 <- quap(
  alist(
    log_gdp_std~dnorm(mu, sigma),
    mu <- a + b * (rugged_std - 0.215),
    a ~ dnorm(1, 0.1),
    b ~ dnorm(0, 0.3),
    sigma ~ dexp(1)
  ), data = d.A0
)
```
```{r}
precis(m8.1)
```

```{r}
precis(m8.2)
```
the first model to consider if add a new predictor
```{r}
m8.3 <- quap(
  alist(
    log_gdp_std~dnorm(mu, sigma),
    mu <- a + b * (rugged_std - 0.215),
    a ~ dnorm(1, 0.1),
    b ~ dnorm(0, 0.3),
    sigma ~ dexp(1)
  ), data = dd
)
```

```{r}
dd$cid <- ifelse(dd$cont_africa==1, 1, 2)
```
second model use indexing
```{r}
m8.4 <- quap(
  alist(
    log_gdp_std~dnorm(mu, sigma),
    mu <- a[cid] + b * (rugged_std - 0.215),
    a[cid]~dnorm(1, 0.1),
    b ~ dnorm(0, 0.3),
    sigma ~ dexp(1)
  ),data = dd
)
```
```{r}
compare(m8.3, m8.4)
```

```{r}
precis(m8.4, depth = 2)
```
Posterier prediction plot of m8.4
```{r}
mu.A0 <- link(m8.4, data=data.frame(cid=2, rugged_std=rugged_seq))
mu.A1 <- link(m8.4, data=data.frame(cid=1, rugged_std=rugged_seq))

mu.A0.mean <- apply(mu.A0, 2, mean)
mu.A0.ci <- apply(mu.A0, 2, HPDI, prob = 0.97)

mu.A1.mean <- apply(mu.A1, 2, mean)
mu.A1.ci <- apply(mu.A1, 2, HPDI, prob = 0.97)

plot(dd$rugged_std, dd$log_gdp_std)
lines(rugged_seq, mu.A0.mean, col = "red")
lines(rugged_seq, mu.A1.mean, col = "blue")

shade(mu.A0.ci, rugged_seq,  col = col.alpha("red"))
shade(mu.A1.ci, rugged_seq,  col = col.alpha("blue"))


```
Varing slop model
```{r}
m8.5 <- quap(
  alist(
    log_gdp_std ~ dnorm( mu, sigma),
    mu <- a[cid] + b[cid] * (rugged_std - 0.215),
    a[cid] ~ dnorm(1, 0.1),
    b[cid] ~ dnorm(0, 0.3),
    sigma ~ dexp(1)
  ), data = dd
)
```
```{r}
precis(m8.5, depth = 2)
```

```{r}
mu.A0 <- link(m8.5, data=data.frame(cid=2, rugged_std=rugged_seq))
mu.A1 <- link(m8.5, data=data.frame(cid=1, rugged_std=rugged_seq))

mu.A0.mean <- apply(mu.A0, 2, mean)
mu.A0.ci <- apply(mu.A0, 2, HPDI, prob = 0.97)

mu.A1.mean <- apply(mu.A1, 2, mean)
mu.A1.ci <- apply(mu.A1, 2, HPDI, prob = 0.97)

plot(dd$rugged_std, dd$log_gdp_std)

lines(rugged_seq, mu.A0.mean, col = "red")
lines(rugged_seq, mu.A1.mean, col = "blue")

shade(mu.A0.ci, rugged_seq,  col = col.alpha("red"))
shade(mu.A1.ci, rugged_seq,  col = col.alpha("blue"))
```

```{r}
compare(m8.5, m8.4, m8.3)

```

```{r}
waic_list <- WAIC(m8.5, pointwise = T)
```
irrelevant staff, just to see if iteration is possible in base R
```{r}
for (i in c("ha", "lo")) print(i)
```
now to show the association of whether a country is in Africa conditioning on ruggedness
```{r}
muA <- link(m8.5, data = data.frame(cid=1, rugged_std=rugged_seq))
muN <- link(m8.5, data = data.frame(cid=2, rugged_std=rugged_seq))

delta <- muA - muN
```


```{r}
delta.mean <- apply(delta, 2, mean)
delta.CI <- apply(delta, 2, HPDI)

plot(rugged_seq, delta.mean, type="l")
shade(delta.CI, rugged_seq, col=col.alpha("black"))
abline(h=0,lty =2)
```

