---
title: "covariance"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rethinking)
options(mc.cores = parallel::detectCores())
```

### Data Simulation
a simple confounding factor
```{r dgp}
N <- 1000
U <- rnorm(N)
X <- U*0.8 + rnorm(N)
Y <- X*1.2 + U * 0.5 + rnorm(N) 
```

### Goal
the goal is to compare 1) a multiple regression model and 2) a covariance model and see if covariance can substitute the unknown variable 

```{r multiple-regression}
# simple multiple regression should recover the parameter of 1.2
lm(Y~X+U) %>% summary # Umm... close
```
```{r covariance}
cov(X, Y)
```
```{r model without covariance, results = "hidden", cache=TRUE}
m1 <- ulam(
  alist(
    Y ~ dnorm(mu, sigma),
    mu <- X * b,
    sigma ~ dexp(1),
    b ~ dnorm(1, 1)
  ),
  data = list(X = X, Y = Y), cores = 4, chains = 4
)
```
```{r result of simple}
precis(m1)
```

```{r model with covariance, results="hidden", cache=TRUE}
m2 <- ulam(
  alist(
    c(X,Y) ~ multi_normal(c(muX, muY), Rho, Sigma),
    muY <- aY + bXY * X,
    Rho ~ lkj_corr(2),
    Sigma ~ exponential(1),
    aY~dnorm(0,0.2),
    bXY ~dnorm(0,0.5),
    vector[1000]:muX ~ dnorm(0, 0.2)
  ),
  data = list(X = X, Y = Y), cores = 4, chains = 4
)
```
```{r}
precis(m2, depth = 3)
```
the result showed the bXY to be 0.67 [-0.04, 1.29] weaker than the real figure of 1.2 and have uncertain sign. given the muX was modeled, it must have eaten up lots of variability of Y.  

```{r model with covariance2, results="hidden", cache=TRUE}
m3 <- ulam(
  alist(
    c(X,Y) ~ multi_normal(c(X, muY), Rho, Sigma),
    muY <- aY + bXY * X,
    Rho ~ lkj_corr(2),
    Sigma ~ exponential(1),
    aY~dnorm(0,0.2),
    bXY ~dnorm(0,0.5)
  ),
  data = list(X = X, Y = Y), cores = 4, chains = 4
)
```

```{r}
precis(m3, depth = 3)
```
Well this seems to be entirely off the mark too. bXY was estimated between [1.40, 1.45], actually the same as a naive linear regression - but this ran for more than 20 minutes! Surely the covariance modelling cannot produce miracles
```{r}
lm(Y ~ X) %>% summary()
```

