---
title: "for kiki"
author: "paul"
date: "3/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
Data
```{r}
df <- tibble::tribble(
  ~sales, ~prom, ~adv, ~index, ~q1, ~q2, ~q3, ~q4,
  504.72,  15.6,  30L,   100L,  1L,  0L,  0L,  0L,
  406.59,  22.2,  36L,   102L,  0L,  1L,  0L,  0L,
  398.55,     0,  45L,   104L,  0L,  0L,  1L,  0L,
  587.76,     0,  57L,   104L,  0L,  0L,  0L,  1L,
  598.92,     0,  39L,   104L,  1L,  0L,  0L,  0L,
  703.62,  31.8,  21L,   100L,  0L,  1L,  0L,  0L,
  387.24,  21.3,  12L,    98L,  0L,  0L,  1L,  0L,
  365.67,   3.9,   6L,    96L,  0L,  0L,  0L,  1L,
  388.71,     0,   6L,    98L,  1L,  0L,  0L,  0L,
  372.96,   8.4,  30L,   103L,  0L,  1L,  0L,  0L,
   603.3,  45.3,  30L,   105L,  0L,  0L,  1L,  0L,
  614.73,  50.1,  33L,   107L,  0L,  0L,  0L,  1L,
  484.38,  39.6,   6L,   107L,  1L,  0L,  0L,  0L,
  227.76,   4.2,  33L,   107L,  0L,  1L,  0L,  0L,
  329.13,     0,   6L,   108L,  0L,  0L,  1L,  0L,
  308.25,     0,   3L,   105L,  0L,  0L,  0L,  1L,
  433.86,     0,  45L,   103L,  1L,  0L,  0L,  0L,
  514.98,  13.8,  48L,   108L,  0L,  1L,  0L,  0L,
   404.7,  17.7,   0L,   110L,  0L,  0L,  1L,  0L,
  245.43,     0,  15L,   112L,  0L,  0L,  0L,  1L,
   433.2,  17.4,   9L,   113L,  1L,  0L,  0L,  0L,
  627.24,  37.8,  54L,   112L,  0L,  1L,  0L,  0L,
  647.61,  42.3,  36L,   113L,  0L,  0L,  1L,  0L,
  342.81,  11.4,  39L,   114L,  0L,  0L,  0L,  1L
  )

```
Create lagged variables and cold season
```{r}
df <- df %>% 
  mutate(lagged_prom1 = lag(prom, 1),
         lagged_prom2 = lag(prom, 2),
         lagged_adv1 = lag(adv, 1),
         lagged_adv2 = lag(adv, 2),
         cold = case_when(
           (q1 == 1)|(q2==1) ~ 1,
           TRUE ~ 0
         )
         )
```
Prom's total effects on sales
```{r}
lm(data = df, sales ~ prom) %>% summary

```
Adv's total effects on sales
```{r}
lm(data = df, sales ~ adv) %>% summary
```
Considering lagged prom, 1 dollar of prom result in  about 3 dollar sales increase across periods 
```{r}
lm(data = df, sales ~ prom + lagged_prom1) %>% summary
```
adv is more effective to later quarter
```{r}
lm(data = df, sales ~ lagged_adv1) %>% summary
```
Full model explained over 70% of sales variations
```{r}
lm(data = df, sales ~ lagged_adv1 + lagged_prom1 + prom) %>% summary
```
Index have no predictive power to sales

```{r}
lm(data = df, sales ~ index) %>% summary
```
Correlation very close to zero
```{r}
cor(df$sales, df$index)
```
Impact of quarters not significant
```{r}
lm(data = df, sales ~ q1 + q2 + q3) %>% summary
```
Impact of quarters not significant even conditioned on all other factors
```{r}
lm(data = df, sales ~ q1 + q2 + q3 + lagged_adv1 + lagged_prom1 + prom) %>% summary
```
Divide quarters into seasons, no impact
```{r}
lm(data = df, sales~cold) %>% summary
```
Adopting full model, still no impact
```{r}

lm(data = df, sales ~ cold + lagged_adv1 + lagged_prom1 + prom) %>% summary
```

