---
title: "R Notebook"
output: html_notebook
---
# Week 1 Homework question 3
Construct the problem step by step with testing
## First generate experiment data, the aim is to find out how many trials is powerful enough to have a more precise credible interval under grid approximation
```{r setup}
library(tidyverse)
library(rethinking)
```


```{r}
experiment_result <- function(trials){
  rbinom(1, trials, prob = 0.7)
}

experiment_data <- function(trials){
  c(result = experiment_result(trials), n = trials)
}
```
test experiment data generation
```{r}
test <- lapply(2^(1:4), experiment_data)
```
Constrcut the grid and prior
```{r}
p_grid <- seq(0, 1, length.out=1000)
prior <- c(rep(0,500), rep(1:500))
prior <- prior / sum(prior)
```
Now the likelihood function, i.e. how likely is the data if given parameter
```{r}
likelihood <- function(data, parameter) {
  dbinom(data["result"], data["n"], prob = parameter)
}
```
```{r}
test <- mapply(likelihood, test, 0.7)
```
generate posterior
```{r}
get_spread <- function(num_trials) { 
  data <- experiment_data(num_trials)
  posterior <- prior * likelihood(data, p_grid)
  posterior <- posterior / sum(posterior)
  c(n = num_trials, calc_spread(posterior))
}

calc_spread <- function(posterior){
  samples <- sample(p_grid, 1e4, replace = T, prob = posterior)
  upper_value <- quantile(samples, 0.995)
  lower_value <- quantile(samples, 0.005)
  
  names(upper_value) <- NULL
  names(lower_value) <- NULL
  c(upper = upper_value, lower =  lower_value, spread = upper_value - lower_value)
}
```
```{r}
result <- data.frame(t(sapply(2^(1:12), get_spread)))
```
```{r}
result %>% 
  ggplot(aes(x = n, y= spread))+
  geom_line()+
  geom_errorbar(aes(ymax = upper, ymin = lower))
```
```{r}
# data.frame() %>% 
  data.table(n = seq(2,2.5, length.out = 5) * 1000)[, as.list(get_spread(n)), by = n]
```
```{r}
data.table(n=2^(1:12))
```

```{r}
data(cherry_blossoms)
d <- cherry_blossoms
precis(d)
```

```{r}
B <- methods(class = class(d))
```

```{r}
WAIC(d)
```
```{r}
B <- methods(precis)
```

```{r}
attr(B, "info")
```

