---
title: "R Notebook"
output: html_notebook
---

# Week 1 homework

## 1 Grid approximation

flat prior 8 water in 15 toss
```{r}
prior <- seq(from=0, to=1, length.out = 1000) # I shouldn't think this is the prior, in fact this is the probability grid
```
```{r}
likelyhood <- function(probability){
  dbinom(8, size = 15, prob = probability)
}
```
```{r}
posterior <- likelyhood(prior)
```
```{r}
posterior <- posterior / sum(posterior)
```
### here I think I missed the part where I shouldn't sample from the posterior, the posterior is just a probability function
```{r}
plot(prior,posterior, type="l")
```
## the answer have some additional steps
```{r}
p_grid <- seq(0, 1, length.out = 1000)
prior  <- rep(1, 1000)
likelyhoods <- likelyhood(p_grid)
posterior <- prior * likelyhoods
posterior <- posterior / sum(posterior)
samples1 <- sample(p_grid, size= 1e4, replace=T, prob = posterior)
```
```{r}
plot(density(samples1))
```
```{r}
summary(samples)
```
## 2 replace flat prior with a conviction that the earth's surface is mostly covered by water
```{r}
p_grid <- seq(0, 1, length.out = 1000)
prior <- c(rep(0, 500), rep(1, 500))
prior <- prior / sum(prior)
likelihoods <- likelyhood(p_grid)
posterior <- prior * likelihoods
posterior <- posterior / sum(posterior)
samples2 <- sample(p_grid, 1e4, replace = T, prob = posterior)

plot(density(samples1))
abline(v=0.7)
plot(density(samples2))
abline(v=0.7)

```
```{r}
quantile(samples1, c(.95, .995))
quantile(samples2, c(.95, .995))
```
```{r}
library(rethinking)
```
```{r}
dens(samples1, xlab = "p", xlim=c(0,1), ylim=c(0,6), show.HPDI = T)
dens(samples2, add = T, lty=2, show.HPDI = T)
abline(v=0.7)
```

## 3 Power of sample numbers
Build functions step-by-step
observed data generation, with p = 0.7
```{r}
observed <- function(trials){
  rbinom(1, size = trials, prob = 0.7)
}
```

```{r}

likelihood <- function(p, observed_data, trials){
  dbinom(observed_data, trials, prob = p)
}

```


```{r}
trial = 32

p_grad <- seq(0,1, length.out = 1000)

prior <- c(rep(0, 500), rep(1,500))
prior <- prior / sum(prior)
observed_data <- observed(trial)

likelihoods <- likelihood(p_grad, observed_data, trial)

posterior <- prior * likelihoods

posterior <- posterior / sum(posterior)

samples = sample(p_grad, 1e4, replace = T, prob = posterior)


```


test the overall data generation
```{r}
lapply(2^(0:3), observed)
```


